---
title: "Exploration of CLEANED QT json file"
author: "Peter Steward"
date: "2024-07-01"
output:
  pdf_document:
    toc: true
    toc_depth: 6
  html_document:
    toc: true
    toc_float: true
    toc_depth: 6
    df_print: paged
---

```{r setup,echo=F}
knitr::opts_chunk$set(echo = TRUE)

# Load packages #####
if (!require("pacman")) {
  install.packages("pacman")
  require(pacman)
}

if (!tinytex::is_tinytex()) {
  p_load(tinytex)
  tinytex::install_tinytex()
}

# Use p_load to install if not present and load the packages
p_load(jsonlite,data.table,rmarkdown,knitr)
```

## Cleaned Data
### Input data structure

The aim of this markdown document is to summarize the structure of the .json object provided from the QT UI.

#### Load the json file into R
```{r load data,echo=T}
# Read in example json file #####
json_data <- fromJSON("qt_example.json")

# Summarize level 1 of the list
json_names<-data.table(field_name=names(json_data),
                          class=sapply(json_data,base::class),
                          length=sapply(json_data,length))

kable(json_names, caption = "Summary of Level 1 of the JSON Object")
```

#### Simple fields
```{r simple fields,echo=T}
simple_fields<-json_names[class!="data.frame" & length==1,field_name]
simple_fields<-json_data[json_names[class!="data.frame" & length==1,field_name]]
simple_fields<-data.table(field=names(simple_fields),
                          value=unlist(simple_fields),
                          class=json_names[class!="data.frame" & length==1,class])
```

```{r simple feilds 2,echo=F}
kable(simple_fields, caption = "Summary of simple fields containing a single value")
```

#### Livestock
The `json_data$livestock` of the json list is a data.table that contains information about livestock"]] herd structure and management, manure management and productivity.

```{r livestock 0 ,echo=F}
livestock<-t(json_data$livestock) # Transpose
colnames(livestock)<-paste0("livetype_code_",json_data$livestock$livetype_code)
kable(livestock,caption="Transposed livestock herd data t(json_data$livestock)")
```

The `livestock` are linked to the `feed_basket$feeds` tables by the `livetype_code` keyfield.
```{r livestock 1 ,echo=T}
json_data$livestock[,c("livetype_code","livetype_desc")] # Codes in the feed_items table
lapply(json_data$feed_basket$feeds,"[[","livestock") # Codes in the feed_basket tables
```

#### Feed items
The `json_data$feed_items` level of the json list is a data.table that contains information about the production management and context of feed production.   

```{r feed items 0 ,echo=F}
feed_items<-t(json_data$feed_items) # Transpose
colnames(feed_items)<-paste0("feed_item_code_",json_data$feed_items$feed_item_code)
kable(feed_items,caption="Transposed feed item data t(json_data$feed_items)")
```

The `feed_items` are linked to the `feed_basket$feeds` tables by the `feed_item_code` field.
```{r feed items 1 ,echo=T}
json_data$feed_items[,c("feed_item_code","feed_item_name")] # Codes in the feed_items table
lapply(json_data$feed_basket$feeds,"[[","feed_item_code") # Codes in the feed_basket tables
```

Other keyfields include `feed_type_code` and `land_cover`:
```{r feed items 2 ,echo=T}
json_data$feed_items[,c("feed_type_code","feed_type_name","land_cover","land_cover_desc")]
```

#### Fertilizer
The `json_data$fertilizer` level of the json list is a data.table that contains information about the production management and context of feed production.   

```{r fertilizer 1,echo=F}
kable(json_data$fertilizer,caption="Fertilizer data input table  (json_data$fertilizer)")
```

The values in `fertilizer_code` field do not appear directly correspond to any fields in the feed basket or feed item tables. However columns with similar names do appear in the field item tables

```{r fertilizer 2,echo=T}
fertilizers<-json_data$fertilizer$fertilizer_desc
fi_cols<-colnames(json_data$feed_items)

# Reformat fertilizer names to match column names in the feed_items table
(fertilizers<-gsub(" ","_",tolower(json_data$fertilizer$fertilizer_desc)))

# Find matching columns
fi_cols[fi_cols %in% fertilizers]
```

#### Seasons
The `json_data$season` level of the input data is 2-column table that records the length of each season (adding up to 365 days max).
```{r seasons 1,echo=F}
kable(json_data$seasons,caption="Feed items data input table (json_data$season)")
```

The `season_name` field is the key field that links to the feed basket.
```{r seasons 2 ,echo=T}
json_data$feed_basket$season_name
```

#### Feed basket
The `json_data$feed_basket` level of the input data contains a futher `r length(json_data$feed_basket)` list levels called 
`r paste(names(json_data$feed_basket),collapse=" and ")`. These sub-levels appear to be the same length: `r paste0(names(json_data$feed_basket)," = ", sapply(json_data$feed_basket,length))`.

##### Seasons 
The `json_data$feed_basket$season` object is simple, being a vector containing the names of the seasons:
```{r feed basket seasons,echo=F}
kable(json_data$feed_basket$season_name,caption="Feed basket/season data input table  (json_data$feed_basket$season)")
```

##### Feed Basket 
The `json_data$feed_basket$feed` object is the most structurally complex element of the input json data containing several levels of nesting.
```{r feed basket feeds 1 ,echo=T}
str(json_data$feed_basket$feeds)
```

Each feed basket table (e.g., `json_data$feed_basket$feeds[[1]]`) is constructed using key fields that describe feed items found in the `json_data$feed_items table`, specifically `feed_item_code` and `feed_type_code.` The feeds are then allocated to the herd elements described in the `json_data$livestock table`.

```{r feed basket feeds 0 ,echo=F}
kable(json_data$feed_basket$feeds[[1]],caption="Feed basket/feed data input table tablejson_data$feed_basket$feeds[[1]]")
```

The livestock field in the feed basket table contains a list with `r length(json_data$feed_basket$feeds[[1]]$livestock)` elements, indicating a one-to-many relationship between the diet item and elements of the livestock herd. Each feed item represented by a row in the feed basket table is allocated to different herd elements, with the `livetype_code` field serving as the key field linking the two tables.

```{r feed basket feeds 2 ,echo=F}
kable(json_data$feed_basket$feeds[[1]]$livestock,caption="Feed basket/feed/livestock data input table tablejson_data$feed_basket$feeds[[1]]$livestock")
```

The first element of the livestock list contains no further nesting:
```{r feed basket feeds 3 ,echo=T}
str(json_data$feed_basket$feeds[[1]]$livestock[[1]])
```

### R-project datasets
```{r list rda files}
rda_files<-list.files(".rda")
```

#### GHG parameters
The ghg para object (`cleaned/data/ghg_para.rda`) is a list of tables that appear to refer to IPCC equations, livestock parameters and fertilizers.
```{r ghg_para 1}
load("ghg_para.rda")

# Summarize level 1 of the list
ghg_names<-data.table(field_name=names(ghg_para),
                          dim=sapply(ghg_para,dim))

kable(ghg_names,caption="Tables within ghg_names list")
```
The exception is `ghg_para$table_5.11` which is a list:
```{r ghg_para 2}
ghg_para$table_5.11
```

##### Livestock_parameters
Note there is an invalid character in this table:
```{r ghg_para 3a}
ghg_para$livestock_parameters$`IPCC Category - methane emissions enteric fermentation - Tier 2`[11]<-"Other Cattle or Buffalo - grazing"
```


```{r ghg_para 3}
kable(ghg_para$livestock_parameters,caption="ghg_para$livestock_parameters")
```

##### Fertilizer table
```{r ghg_para 4}
kable(ghg_para$fertilizer_table,caption="ghg_para$fertilizer_table")
```

##### IPCC Tables

###### Table_10.12
```{r ghg_para Table_10.12}
kable(ghg_para$Table_10.12)
```

###### table_10.17
```{r ghg_para table_10.17}
kable(ghg_para$table_10.17)
```

###### table_10.19
```{r ghg_para table_10.19}
kable(ghg_para$table_10.19)
```

###### table_10.21
```{r ghg_para table_10.21}
kable(ghg_para$table_10.21)
```

###### table_10.22
```{r ghg_para table_10.22}
kable(ghg_para$table_10.22)
```

###### table_10A_9
```{r ghg_para table_10A_9}
kable(ghg_para$Table_10.12)
```

###### table_11.1_&_table_11.3
```{r ghg_para table_11.1_&_table_11.3}
kable(ghg_para$`table_11.1_&_table_11.3`)
```

###### table_2.5
```{r ghg_para table_2.5}
kable(ghg_para$table_2.5)
```

###### table_5.12
```{r ghg_para table_5.12}
kable(ghg_para$table_5.12)
```

###### table_5.13
```{r ghg_para table_5.13}
kable(ghg_para$table_5.13)
```

###### table_5.14
```{r ghg_para table_5.14}
kable(ghg_para$table_5.14)
```

#### Stock change
The stock change object (`cleaned/data/stock_change_para.rda`) is a series of nested lists that describe: 1) landuse, management, and input factor_variables for grassland; and 2) landuse, tillage, and input input factor_variables for cropland.

```{r stock change 1}
load("stock_change_para.rda")
str(stock_change_para)
```

##### Grass factor variables
**Grassland: landuse**
```{r stock change grass 1}
x<-unlist(stock_change_para$grassland$landuse[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```
**Grassland: management**
```{r stock change grass 2}
x<-unlist(stock_change_para$grassland$management[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))

```
**Grassland: input**
```{r stock change grass 3}
x<-unlist(stock_change_para$grassland$input[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```

##### Cropland factor variables
**Cropland: landuse**
```{r stock change cropland 1}
x<-unlist(stock_change_para$cropland$landuse[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```
**Cropland: tillage**
```{r stock change cropland 2}
x<-unlist(stock_change_para$cropland$tillage[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```
**Cropland: input**
```{r stock change cropland 3}
x<-unlist(stock_change_para$cropland$input[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```

#### mufindi
The mufindi object (`cleaned/data/mufindi.rda`) appears to be an input dataset similar to the `qt_example.json` file.
```{r mufindi 1}
load("mufindi.rda")
str(mufindi)
```
