---
title: "Exploration of CLEANED QT json file"
author: "Peter Steward"
date: "2024-07-01"
output:
  pdf_document:
    toc: true
    toc_depth: 6
  html_document:
    toc: true
    toc_float: true
    toc_depth: 6
    df_print: paged
---

```{r setup,echo=F}
knitr::opts_chunk$set(echo = TRUE)

# Load packages #####
if (!require("pacman")) {
  install.packages("pacman")
  require(pacman)
}

if (!tinytex::is_tinytex()) {
  p_load(tinytex)
  tinytex::install_tinytex()
}


# Use p_load to install if not present and load the packages
p_load(jsonlite,data.table,rmarkdown,knitr)
```

## Cleaned Data
### Input data structure

The aim of this markdown document is to summarize the structure of the .json object provided from the QT UI.

#### Load the json file into R
```{r load data,echo=T}
# Read in example json file #####
json_data <- fromJSON("qt_example.json")

# Summarize level 1 of the list
json_names<-data.table(field_name=names(json_data),
                          class=sapply(json_data,base::class),
                          length=sapply(json_data,length))

kable(json_names, caption = "Summary of Level 1 of the JSON Object")
```

#### Simple fields
```{r simple fields,echo=T}
simple_fields<-json_names[class!="data.frame" & length==1,field_name]
simple_fields<-json_data[json_names[class!="data.frame" & length==1,field_name]]
simple_fields<-data.table(field=names(simple_fields),
                          value=unlist(simple_fields),
                          class=json_names[class!="data.frame" & length==1,class])
kable(simple_fields, caption = "Summary of simple fields containing a single value")
```

#### Livestock
The livestock level of the json list is a data.table that contains information about livestock herd structure and management, manure management and productivity.

Here is the data transposed into long form so it is easier to fit onto the page.
```{r livestock 0 ,echo=T}
livestock<-t(json_data$livestock) # Transpose
colnames(livestock)<-paste0("livetype_code_",json_data$livestock$livetype_code)
kable(livestock,caption="Transposed livestock herd data")
```

The `livestock` are linked to the `feed_basket$feeds` tables by the `livetype_code` keyfield.
```{r livestock 1 ,echo=T}
json_data$livestock[,c("livetype_code","livetype_desc")] # Codes in the feed_items table
lapply(json_data$feed_basket$feeds,"[[","livestock") # Codes in the feed_basket tables
```

#### Feed items
The feed_items level of the json list is a data.table that contains information about the production management and context of feed production.   

Here is the data transposed into long form so it is easier to fit onto the page.
```{r feed items 0 ,echo=T}
feed_items<-t(json_data$feed_items) # Transpose
colnames(feed_items)<-paste0("feed_item_code_",json_data$feed_items$feed_item_code)
kable(feed_items,caption="Transposed feed item data")
```

The `feed_items` are linked to the `feed_basket$feeds` tables by the `feed_item_code` field.
```{r feed items 1 ,echo=T}
json_data$feed_items[,c("feed_item_code","feed_item_name")] # Codes in the feed_items table
lapply(json_data$feed_basket$feeds,"[[","feed_item_code") # Codes in the feed_basket tables
```

Other keyfields include `feed_type_code` and `land_cover`:
```{r feed items 2 ,echo=T}
json_data$feed_items[,c("feed_type_code","feed_type_name","land_cover","land_cover_desc")]
```

#### Fertilizer
The fertilizer level of the json list is a data.table that contains information about the production management and context of feed production.   

```{r fertilizer 1,echo=T}
json_data$fertilizer
```

The values in `fertilizer_code` field do not appear directly correspond to any fields in the feed basket or feed item tables. However columns with similar names do appear in the field item tables

```{r fertilizer 2,echo=T}
fertilizers<-json_data$fertilizer$fertilizer_desc
fi_cols<-colnames(json_data$feed_items)

# Reformat fertilizer names to match column names in the feed_items table
(fertilizers<-gsub(" ","_",tolower(json_data$fertilizer$fertilizer_desc)))

# Find matching columns
fi_cols[fi_cols %in% fertilizers]
```

#### Seasons
The seasons level is simple table that records the length of each season (adding up to 365 days max).
```{r seasons 1,echo=T}
json_data$seasons
```

The `season_name` field is the key field that links to the feed basket.
```{r seasons 2 ,echo=T}
json_data$feed_basket$season_name
```

#### Feed basket
The feed basket data  contains `r length(json_data$feed_basket)` levels called 
`r paste(names(json_data$feed_basket),collapse=" and ")`. These levels appear to be the same length: `r paste0(names(json_data$feed_basket)," = ", sapply(json_data$feed_basket,length))`.

##### Seasons 
The season object is simple, being a vector containing the names of the seasons:
```{r feed basket seasons,echo=T}
json_data$feed_basket$season_name
```

##### Feed Basket 
The feed basket object is the most structurally complex element of the input json data containing several levels of nesting.
```{r feed basket feeds 1 ,echo=T}
str(json_data$feed_basket$feeds)
```

Each feed basket table (e.g., `json_data$feed_basket$feeds[[1]]`) is constructed using key fields that describe feed items found in the `json_data$feed_items table`, specifically `feed_item_code` and `feed_type_code.` The feeds are then allocated to the herd elements described in the `json_data$livestock table`.

```{r feed basket feeds 0 ,echo=T}
json_data$feed_basket$feeds[[1]]
```

The livestock field in the feed basket table contains a list with `r length(json_data$feed_basket$feeds[[1]]$livestock)` elements, indicating a one-to-many relationship between the diet item and elements of the livestock herd. Each feed item represented by a row in the feed basket table is allocated to different herd elements, with the `livetype_code` field serving as the key field linking the two tables.

```{r feed basket feeds 2 ,echo=T}
json_data$feed_basket$feeds[[1]]$livestock
```

The first element of the livestock list contains no further nesting:
```{r feed basket feeds 3 ,echo=T}
str(json_data$feed_basket$feeds[[1]]$livestock[[1]])
```

### R-project datasets
```{r list rda files}
rda_files<-list.files(".rda")
```

#### GHG parameters
The ghg para object (`cleaned/data/ghg_para.rda`) is a list of tables that appear to refer to IPCC equations, livestock parameters and fertilizers.
```{r ghg_para 1}
load("ghg_para.rda")

# Summarize level 1 of the list
ghg_names<-data.table(field_name=names(ghg_para),
                          dim=sapply(ghg_para,dim))

kable(ghg_names,caption="Tables within ghg_names list")
```
The exception is `ghg_para$table_5.11` which is a list:
```{r ghg_para 2}
ghg_para$table_5.11
```

##### Livestock_parameters
```{r ghg_para 3}
ghg_para$livestock_parameters
```

##### Fertilizer table
```{r ghg_para 4}
ghg_para$fertilizer_table
```

##### IPCC Tables

###### Table_10.12
```{r ghg_para Table_10.12}
kable(ghg_para$Table_10.12)
```

###### table_10.17
```{r ghg_para table_10.17}
kable(ghg_para$table_10.17)
```

###### table_10.19
```{r ghg_para table_10.19}
kable(ghg_para$table_10.19)
```

###### table_10.21
```{r ghg_para table_10.21}
kable(ghg_para$table_10.21)
```

###### table_10.22
```{r ghg_para table_10.22}
kable(ghg_para$table_10.22)
```

###### table_10A_9
```{r ghg_para table_10A_9}
kable(ghg_para$Table_10.12)
```

###### table_11.1_&_table_11.3
```{r ghg_para table_11.1_&_table_11.3}
kable(ghg_para$`table_11.1_&_table_11.3`)
```

###### table_2.5
```{r ghg_para table_2.5}
kable(ghg_para$table_2.5)
```

###### table_5.12
```{r ghg_para table_5.12}
kable(ghg_para$table_5.12)
```

###### table_5.13
```{r ghg_para table_5.13}
kable(ghg_para$table_5.13)
```

###### table_5.14
```{r ghg_para table_5.14}
kable(ghg_para$table_5.14)
```

#### Stock change
The stock change object (`cleaned/data/stock_change_para.rda`) is a series of nested lists that describe: 1) landuse, management, and input factor_variables for grassland; and 2) landuse, tillage, and input input factor_variables for cropland.

```{r stock change 1}
load("stock_change_para.rda")
str(stock_change_para)
```

##### Grass factor variables
**Grassland: landuse**
```{r stock change grass 1}
x<-unlist(stock_change_para$grassland$landuse[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```
**Grassland: management**
```{r stock change grass 2}
x<-unlist(stock_change_para$grassland$management[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))

```
**Grassland: input**
```{r stock change grass 3}
x<-unlist(stock_change_para$grassland$input[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```

##### Cropland factor variables
**Cropland: landuse**
```{r stock change cropland 1}
x<-unlist(stock_change_para$cropland$landuse[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```
**Cropland: tillage**
```{r stock change cropland 2}
x<-unlist(stock_change_para$cropland$tillage[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```
**Cropland: input**
```{r stock change cropland 3}
x<-unlist(stock_change_para$cropland$input[[1]]$factor_variables)
kable(data.frame(variable=names(x),value=as.numeric(x)))
```

#### mufindi
The mufindi object (`cleaned/data/mufindi.rda`) appears to be an input dataset similar to the `qt_example.json` file.
```{r mufindi 1}
load("mufindi.rda")
str(mufindi)
```
